---
title: "In-class Ex4"
author: "Jayexx Low"
date: "May 4, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

### Installing & loading required libraries

The code chunk below installs and launches the tidyverse, ggdist, ggridges, colourspace & ggthemes packages into R environment

```{r}
pacman:: p_load(ggstatsplot, tidyverse)
```

### Importing the data

The code chunk below imports exam_data.csv into R environment by using read_csv() function of readr package, which is part of the tidyverse package.

```{r}
exam <- read_csv("data/Exam_data.csv")
``` 

In the code chunk below, gghistostats() is used to to build an visual of one-sample test on English scores.

```{R}
set.seed(1234)

p <- gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  bin.args = list(color = "black",
                  fill = "grey50",
                  alpha = 0.7),
                  normal.curve = TRUE,
                  normal.curve.args = list(linewidth = 0.5),
  xlab = "English scores"
)

p
```

```{R}
extract_stats(p)
```

```{R}
ggdotplotstats(
  data = exam,
  x = ENGLISH,
  y = CLASS,
  title = "",
  xlab = ""
)
```
Note on the above, the y axis 'CLASS' categorical variable is arranged according to its x axis values rather than alphabetical order,  eg. 3D is above 3A 

```{R}
exam_long <- exam %>%
  pivot_longer(
    cols = ENGLISH: SCIENCE,
    names_to = "SUBJECT",
    values_to = "SCORES") %>%
  filter(CLASS == "3A")
```




```{R}
ggwithinstats(
  data = filter(exam_long,
                SUBJECT %in%
                  c("MATHS", "SCIENCE")),
  x = SUBJECT, 
  y = SCORES,
  type="p"
)
```

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y= ENGLISH,
  marginal = TRUE,
  label.var = ID,
  label.expression = ENGLISH > 90 & MATHS >90
)
```







