---
title: "In-class Ex2"
author: "Jayexx Low"
date: "April 20, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

### Install & Load Libraries

The code chunk below installs and launches the tidyverse, ggdist, ggridges, colourspace & ggthemes packages into R environment

```{r}
pacman:: p_load(tidyverse, ggdist, ggridges,
                colourspace, ggthemes)
```

### Importing the data

The code chunk below imports exam_data.csv into R environment by using read_csv() function of readr package, which is part of the tidyverse package.

```{r}
exam <- read_csv("data/Exam_data.csv")
```

## Histogram

```{R}
ggplot(data=exam, 
       aes(x = ENGLISH)) +
  geom_histogram(
    color = "#1696d2",
  )
```

## Probability Density plot

```{R}
ggplot(data=exam, 
       aes(x = ENGLISH)) +
  geom_density(
    color = "#1696d2",
    adjust = .65,
    alpha = .6
  )           
```

## Alternative design


ggplot(data=exam_df, 
       aes(x = ENGLISH)) +
  geom_density(
    color = "#1696d2",
    adjust = .65,
    alpha = .6
  ) +
  stat_function(
    fun = dnorm, args = list(mean = mean_eng,
                             sd = std_eng),
    col = "grey30",
    size = .8
  ) +
  geom_vline(
    aes(xintercept = mean_eng),
    colour="4d5887",
    linewidth = .6,
    linetype = "dashed"
    ) +
  annotate(geom = "text",
           x = mean_eng - 8,
           y = 0.04,
           label = paste)

## Visualising Distribution with Ridgeline Plot

```{R}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```






## Visualising Distribution with Raincloud Plot

